---
title: "tidy data"
output: github_document
---
```{r}
library(tidyverse)
```

## PULSE data


```{r}
pulse_df = 
  haven::read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(bdi_score_bl:bdi_score_12m,
               names_to = "visit",
               values_to = "bdi_scores", 
               names_prefix = "bdi_score_"
               ) |>
      mutate(
        visit = replace(visit, visit == "bl", "00m")
      )
  
```

## learning assessment 

In the litters data, the variables gd0_weight and gd18_weight give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only litter_number and these columns; produces new variables gd and weight; and makes gd a numeric variable taking values 0 and 18 (for the last part, you might want to use recode …). 

import . kentgthen the litters dataset
```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") |>
  janitor::clean_names() |>
  select(litter_number, gd0_weight, gd18_weight) |>
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight",
    names_prefix = "gd"
  ) |> 
  mutate (
    gd= case_match (
      gd, 
      "0_weight" ~ 0, 
      "18_weight" ~ 18 # any time see 18 weight, replace it with 18
    )
  )

```

## LoTR

import LoTR wirds data

```{r}
fellowship_df = 
  readxl::read_excel('data/LotR_Words.xlsx', range = "B3:D6") |>
  mutate(movie = "fellowship")


two_towers_df = 
  readxl::read_excel('data/LotR_Words.xlsx', range = "F3:H6") |>
  mutate(movie = "two towers")

return_of_the_king_df = 
  readxl::read_excel('data/LotR_Words.xlsx', range = "J3:L6") |>
  mutate(movie = "return of the king")


lotr_df = 
  bind_rows (fellowship_df, two_towers_df, return_of_the_king_df) |>
  janitor::clean_names() |>
  pivot_longer (
    male:female, 
    names_to = "gender",
    values_to = "word" ) |>
  relocate(movie)
```














