---
title: "Data Manipulation"
output: github_document
---
```{r}
library(tidyverse)
```

lets import the 'FAS_litters.csv', import FAS litters and pups

```{r}
litters_df = read_csv("./data/FAS_litters.csv")

# clean the name, upper case to lower case...
litters_df = janitor::clean_names(litters_df)
  
pups_data = read_csv(file = "~/Desktop/data_wrangleing_1/data/FAS_pups.csv")

pups_data_df = janitor::clean_names(pups_data)

```

## select

specify the columns you wnat by naming all of them 

```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)

select(litters_df, group, gd0_weight:gd_of_birth) # specify the range of columns to keep 

select(litters_df, group, starts_with("pups")) #select those column start with gd

select(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number) # rename variables

select(litters_df, -litter_number) # remove variables 
```

learning assessment 

```{r}
select(pups_data_df, litter_number, sex, pd_ears)
```

## filter

get rid of rows using filter()
filter() subsets a dataframe by row, select() returns a subset of the columns. 


```{r, eval = FALSE}
filter(litters_df, group == "Con7")

filter(litters_df, group == "Mod8")

filter(litters_df, group != "Con7")

filter(litters_df, gd0_weight > 20)

filter(litters_df, gd0_weight <= 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, !(group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20)
```

learning check 

```{r}
filter(pups_data_df, sex==1)
```


what about missing rows

```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```

## Mutate

change variable names (modify existing variable names)
new variables can be functions of old variables
new variables appear at the end of the dataset in the order that they are created
you can overwrite old variables
you can create new variables and immediatley change it. 

```{r}
mutate(litters_df,  wt_gain = gd18_weight - gd0_weight)

mutate(litters_df,  group = str_to_lower(group))

mutate(litters_df,  
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group)
)
```

## Arrange

arrange rows in data according to the values in columns. 
```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, group, gd0_weight)
```

## pipes
|> is used for nested functions 

ccddiii??
```{r}
litters_df =
  read.csv("data/FAS_litters.csv") |>
  janitor::clean_names() |>
  select(-starts_with("pups")) |>
  mutate(
    group = str_to_lower(group),
    wt_gain = gd18_weight-gd0_weight,
  ) |>
  drop_na(wt_gain) |>
  arrange(group, wt_gain)
```



